Fantasy Football Analysis by Sandeep Mohan
========================================================
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.
library(tidyverse)
library(ggplot2)
library(GGally)
library(ggthemes)
library(markdown)
library(rmarkdown)
library(knitr)
library(RColorBrewer)
library(gridExtra)
library(reshape2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, dataset load}
robust_df = read.csv("~/Desktop/GBPredictions/final_submission/fantasy_football_project/robust_ff_wide.csv")

master_df = read.csv("~/Desktop/GBPredictions/final_submission/fantasy_football_project/master_ff.csv")
```


> **Introduction to the Dataset and Analysis Goals**: TBD


```{r echo=FALSE,message=FALSE, warning=FALSE, correlation heat map}
ggcorr(robust_df)
```

# Univariate Plots Section

>**Age of players**:The overwhelming majority of active NFL players are below 30 
as expected though wide receivers (WR) and quarterbacks (QB) can have longer careers. 
As the top left histogram in the charts below shows Tom Brady of the NE Patriots 
is in a league of his own...

>The dotted line shows the median age of players by position while the dashed line
indicated the 99th percentile.

>Given that the versatile Tom Brady as a QB also needs to sometimes run the ball 
himself, his elite performance thus far is phenomenal

```{r echo=FALSE,message=FALSE, warning=FALSE}

ggplot(robust_df, aes(x=age, color=pos, fill = pos)) +
  geom_histogram(alpha=0.6, position="identity", bins = 20) +
  geom_vline(aes(xintercept = median(robust_df[which(robust_df$pos == pos),]$age, 
                                     na.rm = TRUE)), linetype = "dashed") +
  geom_vline(aes(xintercept = quantile(robust_df[which(robust_df$pos == pos),]$age, 
                                       na.rm = TRUE, 0.99), linetype = "dashed")) +
  labs(x="Age", y="Player Count") + 
  facet_wrap(~pos) +
  theme_tufte(base_size = 10, base_family = "sans", ticks = TRUE)
```

>**Points Expected Per Week in Regular Season**: 
Prior to the draft, its important to know how many points a player can expect to 
net the fantasy team on average that drafts him to the specific position. Experts
provide a season points. Based on the source its not clear that if this calculates
the playoffs but there is no way to tell at the offset which teams are likely to
make the playoffs. So the working assumption is that the projected points are for 
the 16 games in regular season. So the point total has been divided by 16 to generate 
an average expected per game. To the extent that one's players are outperforming
or meeting their expected average, we can calculate the value of the draft pick.

> Dotted line below shows the 95th percentile of players. While targeting in the 
draft, these players will be likely the first drafted. 


```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(robust_df, aes(x=pointspg_w0, color=pos, fill = pos)) +
  geom_histogram(alpha=0.6, position="identity", bins = 20) +
  geom_vline(aes(xintercept = quantile(robust_df[which(robust_df$pos == pos),]$pointspg_w0, 
                                     na.rm = TRUE, 0.95)), linetype = "dotted") +
  labs(x="Average Points Projected over 16 Regular Season Games", y="Player Count") + 
  facet_wrap(~pos) +
  theme_tufte(base_size = 10, base_family = "sans", ticks = TRUE)
```

>**Risk Average**: Similar to Points per game, Injury/Penalty/Missed Games Risk
per week has also been computed below at Season offset. 

>The dotted line below indicates the 75th percentile of risk level per position. 
It looks like this year the risk to RBs are less compared to the Wrs. Needs more 
exploration because that will likely impact the choice of whether to draft WRs or 
RBs in the early rounds of the draft. Also, it looks like QBs are expected to net
more points per game at less risk so probably good idea to get a good QB early

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(robust_df, aes(x=risk_w0, color=pos, fill = pos)) +
  geom_histogram(alpha=0.6, position="identity", bins = 20) +
  geom_vline(aes(xintercept = quantile(robust_df[which(robust_df$pos == pos),]$risk_w0, 
                                     na.rm = TRUE, 0.75)), linetype = "dotted") +
  labs(x="Risk Levels to Players Expected over 16 Regular Season Games", 
       y="Player Count") + 
  facet_wrap(~pos) +
  theme_tufte(base_size = 10, base_family = "sans", ticks = TRUE)
```
```{r}
head(robust_df)
```
```{r}
w_point_proj = subset(robust_df, team != "FA", 
                      select = c(id, name, team, pos, 
                                 pointspg_w0, points_w1, points_w2, points_w3))



w_risk_proj = subset(robust_df, team != "FA", 
                      select = c(id, name, team, pos, 
                                 risk_w0, risk_w1, risk_w2, risk_w3))

w_posecr_proj = subset(robust_df, team != "FA", 
                      select = c(id, name, team, pos, 
                                 pos_ecr_w0, pos_ecr_w1, pos_ecr_w2, pos_ecr_w3))

w_floor_proj = subset(robust_df, team != "FA", 
                      select = c(id, name, team, pos, 
                                 floorpg_w0, floor_w1, floor_w2, floor_w3))

w_ceiling_proj = subset(robust_df, team != "FA", 
                      select = c(id, name, team, pos, 
                                 ceilingpg_w0, ceiling_w1, ceiling_w2, ceiling_w3))
```

```{r}
w_point_proj = melt(w_point_proj, id.vars = c("id", "name", "team", "pos"), 
                    variable.name = "week_proj",
                    value.name = "proj_points")

w_risk_proj = melt(w_risk_proj, id.vars = c("id", "name", "team", "pos"), 
                    variable.name = "week_risk",
                    value.name = "proj_risk")

w_posecr_proj = melt(w_posecr_proj, id.vars = c("id", "name", "team", "pos"), 
                    variable.name = "week_ecr",
                    value.name = "proj_ecr")

w_floor_proj = melt(w_floor_proj, id.vars = c("id", "name", "team", "pos"), 
                    variable.name = "week_floor",
                    value.name = "proj_floor")

w_ceiling_proj = melt(w_ceiling_proj, id.vars = c("id", "name", "team", "pos"), 
                    variable.name = "week_ceiling",
                    value.name = "proj_ceiling")

```
```{r}
str(w_point_proj)
```


```{r echo=FALSE,message=FALSE, warning=FALSE}


ggplot(data = w_point_proj) + 
  geom_boxplot(aes(x = pos, y = proj_points, fill = pos), notch = TRUE) + 
  facet_wrap(~week_proj) + 
  theme_tufte(base_size = 10, base_family = "sans")


```



